### 2014-11-28 matlabcode (1).tex

\documentclass{article}

\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry}
\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
\usepackage{listings}    
\lstset{
        language=Matlab,
        breaklines=true
    }
\pagestyle{fancy}
\lhead{FOVEROS Chess Engine}
\rhead{\thepage}
\cfoot{}
\begin{document}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage
\section{Appendix: The Code}
\newpage

\appendix
\section{Appendix: The Code}
\subsection{Back-end}
\subsubsection{MovementRook.m}
\begin{lstlisting}

function [possiblemoves] = MovementRook(chessboard,piece_colour,p_x,p_y)

%Initialisation values --------------------------------------------------
r_colour = piece_colour(p_x,p_y);
possiblemoves = zeros(8,8);

%This section allows movement in vertical direction -----------------------
i = 1;
while(p_x+i<9)
    if(piece_colour(p_x+i,p_y)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y)~= r_colour && chessboard(p_x+i,p_y)~=0)
        possiblemoves(p_x+i,p_y) = 2;
        break
    end
    possiblemoves(p_x+i,p_y) = 1;
    i = i+1;
end

i = 1;
while(p_x-i>0)
    if(piece_colour(p_x-i,p_y)== r_colour)
        break
    end
     if(piece_colour(p_x-i,p_y)~= r_colour && chessboard(p_x-i,p_y)~=0)
        possiblemoves(p_x-i,p_y) = 2;
        break
    end
    possiblemoves(p_x-i,p_y) = 1;
    i = i+1;
end

%This section allows movement in the horizontal direction
i = 1;
while(p_y+i<9)
    if(piece_colour(p_x,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x,p_y+i)~= r_colour && chessboard(p_x,p_y+i)~=0)
        possiblemoves(p_x,p_y+i) = 2;
        break
    end
    possiblemoves(p_x,p_y+i) = 1;
    i = i+1;
end

i = 1;
while(p_y-i>0)
    if(piece_colour(p_x,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x,p_y-i)~= r_colour && chessboard(p_x,p_y-i)~=0)
        possiblemoves(p_x,p_y-i) = 2;
        break
    end
    possiblemoves(p_x,p_y-i) = 1;
    i = i+1;
end

%-------------------------------------------------------------------------

end
\end{lstlisting}

\subsubsection{MovementQueen.m}
\begin{lstlisting}
function [possiblemoves] = MovementQueen(chessboard,piece_colour,p_x,p_y)

%Initialisation values --------------------------------------------------
possiblemoves = zeros(8,8);
r_colour = piece_colour(p_x,p_y);

%This section allows movement in / direction -----------------------------
i=1;
while(p_x+i<9 && p_y+i<9)
    if(piece_colour(p_x+i,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y+i)~= r_colour && chessboard(p_x+i,p_y+i)~=0)
        possiblemoves(p_x+i,p_y+i) = 2;
        break
    end
    possiblemoves(p_x+i,p_y+i) = 1;
    i = i+1;
end

i=1;
while(p_x-i>0 && p_y-i>0)
    if(piece_colour(p_x-i,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x-i,p_y-i)~= r_colour && chessboard(p_x-i,p_y-i)~=0)
        possiblemoves(p_x-i,p_y-i) = 2;
        break
    end
    possiblemoves(p_x-i,p_y-i) = 1;
    i = i+1;
end

%This section allows movement in the \ direction-------------------------
i=1;
while(p_x+i<9 && p_y-i>0)
    if(piece_colour(p_x+i,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y-i)~= r_colour && chessboard(p_x+i,p_y-i)~=0)
        possiblemoves(p_x+i,p_y-i) = 2;
        break
    end
    possiblemoves(p_x+i,p_y-i) = 1;
    i = i+1;
end

i=1;
while(p_x-i>0 && p_y+i<9)
    if(piece_colour(p_x-i,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x-i,p_y+i)~= r_colour && chessboard(p_x-i,p_y+i)~=0)
        possiblemoves(p_x-i,p_y+i) = 2;
        break
    end
    possiblemoves(p_x-i,p_y+i) = 1;
    i = i+1;
end

%This section allows movement in vertical direction -----------------------------
i = 1;
while(p_x+i<9)
    if(piece_colour(p_x+i,p_y)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y)~= r_colour && chessboard(p_x+i,p_y)~=0)
        possiblemoves(p_x+i,p_y) = 2;
        break
    end
    possiblemoves(p_x+i,p_y) = 1;
    i = i+1;
end

i = 1;
while(p_x-i>0)
    if(piece_colour(p_x-i,p_y)== r_colour)
        break
    end
     if(piece_colour(p_x-i,p_y)~= r_colour && chessboard(p_x-i,p_y)~=0)
        possiblemoves(p_x-i,p_y) = 2;
        break
    end
    possiblemoves(p_x-i,p_y) = 1;
    i = i+1;
end

%This section allows movement in the horizontal direction-------------------------
i = 1;
while(p_y+i<9)
    if(piece_colour(p_x,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x,p_y+i)~= r_colour && chessboard(p_x,p_y+i)~=0)
        possiblemoves(p_x,p_y+i) = 2;
        break
    end
    possiblemoves(p_x,p_y+i) = 1;
    i = i+1;
end

i = 1;
while(p_y-i>0)
    if(piece_colour(p_x,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x,p_y-i)~= r_colour && chessboard(p_x,p_y-i)~=0)
        possiblemoves(p_x,p_y-i) = 2;
        break
    end
    possiblemoves(p_x,p_y-i) = 1;
    i = i+1;
end
%-------------------------------------------------------------------------

end
\end{lstlisting}

\subsubsection{MovementPawn.m}
\begin{lstlisting}
function [possiblemoves] = MovementPawn(chessboard,piece_colour,num_moves,p_x,p_y)

%Initialisation values --------------------------------------------------
r_colour = piece_colour(p_x,p_y);
possiblemoves = zeros(8,8);

%This section allows all movements after checking whether it exceeds the board or not ---------------
switch r_colour
    case 119 %White case
        %En passant-------------------------------------------------------
        if (p_x==4)

            if(p_x-1>0 && p_y-1 >0) %Capture left
                if(piece_colour(p_x,p_y-1)~=r_colour && chessboard(p_x,p_y-1)==1 && num_moves(p_x,p_y-1)==1)
                    possiblemoves(p_x-1,p_y-1) = 3;
                end
            end

            if(p_x-1>0 && p_y+1<9) %Capture right
                if(piece_colour(p_x,p_y+1)~=r_colour && chessboard(p_x,p_y+1)==1 && num_moves(p_x,p_y+1)==1)
                    possiblemoves(p_x-1,p_y+1) = 3;
                end    
            end

        end

        if(p_x-1>0) %Forward movement
            if(chessboard(p_x-1,p_y)==0)
                possiblemoves(p_x-1,p_y) = 1;
            end
        end

        %Initial forward movement
        if(p_x==7 && chessboard(p_x-2,p_y)==0 && chessboard(p_x-1,p_y)==0)
        possiblemoves(p_x-2,p_y) = 1;
        end

        if(p_x-1>0 && p_y-1 >0) %Capture left
            if(piece_colour(p_x-1,p_y-1)~=r_colour && chessboard(p_x-1,p_y-1)~=0)
                possiblemoves(p_x-1,p_y-1) = 2;
                   if(p_x==2) %Capture and pawn promotion
                      possiblemoves(p_x-1,p_y-1) = 5;
                   end
            end
        end

        if(p_x-1>0 && p_y+1<9) %Capture right
            if(piece_colour(p_x-1,p_y+1)~=r_colour && chessboard(p_x-1,p_y+1)~=0)
                possiblemoves(p_x-1,p_y+1) = 2;
                   if(p_x==2) %Capture and pawn promotion
                      possiblemoves(p_x-1,p_y+1) = 5;
                   end
            end
        end

        %Pawn promotion----------------------------------------------------
        if(p_x==2) 
            if(chessboard(p_x-1,p_y)==0)
                possiblemoves(p_x-1,p_y) = 5;
            end
        end

    case 98 %Black Case

        %En passant-------------------------------------------------------
        if (p_x==5)

            if(p_x-1>0 && p_y-1 >0) %Capture left
                if(piece_colour(p_x,p_y-1)~=r_colour && chessboard(p_x,p_y-1)==1 && num_moves(p_x,p_y-1)==1)
                    possiblemoves(p_x+1,p_y-1) = 3;
                end
            end

            if(p_x-1>0 && p_y+1<9) %Capture right
                if(piece_colour(p_x,p_y+1)~=r_colour && chessboard(p_x,p_y+1)==1 && num_moves(p_x,p_y+1)==1)
                    possiblemoves(p_x+1,p_y+1) = 3;
                end    
            end

        end

        if(p_x+1<9) %Forward movement
            if(chessboard(p_x+1,p_y)==0) 
                possiblemoves(p_x+1,p_y) = 1;
            end
        end

        %Initial Forward movement
        if(p_x==2 && chessboard(p_x+2,p_y)==0 && chessboard(p_x+1,p_y)==0)
            possiblemoves(p_x+2,p_y) = 1;
        end

        if(p_x+1<9 && p_y-1>0) %Capture left
            if(piece_colour(p_x+1,p_y-1)~=r_colour && chessboard(p_x+1,p_y-1)~=0)
                possiblemoves(p_x+1,p_y-1) = 2;
                  if(p_x==7) %Capture and pawn promotion
                      possiblemoves(p_x+1,p_y-1) = 5;
                   end
            end
        end

        if(p_x+1<9 && p_y+1<9) %Capture right
            if(piece_colour(p_x+1,p_y+1)~=r_colour && chessboard(p_x+1,p_y+1)~=0)
                possiblemoves(p_x+1,p_y+1) = 2;
                   if(p_x==7) %Capture and pawn promotion
                      possiblemoves(p_x+1,p_y+1) = 5;
                   end
            end   
        end

        %Pawn promotion----------------------------------------------------
        if(p_x==7)
            if(chessboard(p_x+1,p_y)==0) 
                possiblemoves(p_x+1,p_y) = 5;
            end
        end

end
%-------------------------------------------------------------------------
end
\end{lstlisting}

\subsubsection{MovementKnight.m}
\begin{lstlisting}
function [possiblemoves] = MovementKnight(chessboard,piece_colour,p_x,p_y)

%Initialisation values --------------------------------------------------
r_colour = piece_colour(p_x,p_y);
possiblemoves = zeros(8,8);

%This sections allows L shaped movements for knight
if(p_x-2>0 & p_y-1>0)
    if (piece_colour(p_x-2,p_y-1)~= r_colour && chessboard(p_x-2,p_y-1)~=0)
        possiblemoves(p_x-2,p_y-1) = 2;
    elseif (piece_colour(p_x-2,p_y-1)== r_colour)
        ;
    else
        possiblemoves(p_x-2,p_y-1) = 1;
    end
end

if(p_x-2>0 & p_y+1<9)
    if (piece_colour(p_x-2,p_y+1)~= r_colour && chessboard(p_x-2,p_y+1)~=0)
        possiblemoves(p_x-2,p_y+1) = 2;
    elseif (piece_colour(p_x-2,p_y+1)== r_colour)
        ;
    else
        possiblemoves(p_x-2,p_y+1) = 1;
    end
end

if(p_x-1>0 & p_y-2>0)
    if (piece_colour(p_x-1,p_y-2)~= r_colour && chessboard(p_x-1,p_y-2)~=0)
        possiblemoves(p_x-1,p_y-2) = 2;
    elseif (piece_colour(p_x-1,p_y-2)== r_colour)
        ;
    else
        possiblemoves(p_x-1,p_y-2) = 1;
    end
end

if(p_x-1>0 & p_y+2<9)
    if (piece_colour(p_x-1,p_y+2)~= r_colour && chessboard(p_x-1,p_y+2)~=0)
        possiblemoves(p_x-1,p_y+2) = 2;
    elseif (piece_colour(p_x-1,p_y+2)== r_colour)
        ;
    else
        possiblemoves(p_x-1,p_y+2) = 1;
    end
end

if(p_x+1<9 & p_y-2>0)
    if (piece_colour(p_x+1,p_y-2)~= r_colour && chessboard(p_x+1,p_y-2)~=0)
        possiblemoves(p_x+1,p_y-2) = 2;
    elseif (piece_colour(p_x+1,p_y-2)== r_colour)
        ;
    else
        possiblemoves(p_x+1,p_y-2) = 1;
    end
end

if(p_x+1<9 & p_y+2<9)
    if (piece_colour(p_x+1,p_y+2)~= r_colour && chessboard(p_x+1,p_y+2)~=0)
        possiblemoves(p_x+1,p_y+2) = 2;
    elseif (piece_colour(p_x+1,p_y+2)== r_colour)
        ;
    else
        possiblemoves(p_x+1,p_y+2) = 1;
    end
end

if(p_x+2<9 & p_y-1>0)
    if (piece_colour(p_x+2,p_y-1)~= r_colour && chessboard(p_x+2,p_y-1)~=0)
        possiblemoves(p_x+2,p_y-1) = 2;
    elseif (piece_colour(p_x+2,p_y-1)== r_colour)
        ;
    else
        possiblemoves(p_x+2,p_y-1) = 1;
    end
end

if(p_x+2<9 & p_y+1<9)
    if (piece_colour(p_x+2,p_y+1)~= r_colour && chessboard(p_x+2,p_y+1)~=0)
        possiblemoves(p_x+2,p_y+1) = 2;
    elseif (piece_colour(p_x+2,p_y+1)== r_colour)
        ;
    else
        possiblemoves(p_x+2,p_y+1) = 1;
    end
end

%-------------------------------------------------------------------------

end
\end{lstlisting}

\subsubsection{MovementKing.m}
\begin{lstlisting}
function [possiblemoves] = MovementKing(chessboard,piece_colour,num_moves,potential_moves,p_x,p_y)

%Initialisation values --------------------------------------------------
r_colour = piece_colour(p_x,p_y);
possiblemoves = zeros(8,8);

%This section allows all movements after checking whether it exceeds the board or not 
%and ensures that the king is not moving into square that is in check ---------------

%------------------------------------------------------------------------
%                        Movement (8 Directions)
%------------------------------------------------------------------------

    if(p_x+1<9)
        if (piece_colour(p_x+1,p_y)~= r_colour && chessboard(p_x+1,p_y)~=0)
            possiblemoves(p_x+1,p_y) = 2;
        elseif (piece_colour(p_x+1,p_y)== r_colour)
            ;
        else
            possiblemoves(p_x+1,p_y) = 1;
        end
    end

    if(p_x+1<9 && p_y+1<9)
        if (piece_colour(p_x+1,p_y+1)~= r_colour && chessboard(p_x+1,p_y+1)~=0)
            possiblemoves(p_x+1,p_y+1) = 2;
        elseif (piece_colour(p_x+1,p_y+1)== r_colour)
            ;
        else
            possiblemoves(p_x+1,p_y+1) = 1;
        end
    end

    if(p_x+1<9 && p_y-1>0)
        if (piece_colour(p_x+1,p_y-1)~= r_colour && chessboard(p_x+1,p_y-1)~=0)
            possiblemoves(p_x+1,p_y-1) = 2;
        elseif (piece_colour(p_x+1,p_y-1)== r_colour)
            ;
        else
            possiblemoves(p_x+1,p_y-1) = 1;
        end
    end

    if(p_y+1<9)
        if (piece_colour(p_x,p_y+1)~= r_colour && chessboard(p_x,p_y+1)~=0)
            possiblemoves(p_x,p_y+1) = 2;
        elseif (piece_colour(p_x,p_y+1)== r_colour)
            ;
        else
            possiblemoves(p_x,p_y+1) = 1;
        end
    end

    if(p_y-1>0)
        if (piece_colour(p_x,p_y-1)~= r_colour && chessboard(p_x,p_y-1)~=0)
            possiblemoves(p_x,p_y-1) = 2;
        elseif (piece_colour(p_x,p_y-1)== r_colour)
            ;
        else
            possiblemoves(p_x,p_y-1) = 1;
        end
    end

    if(p_x-1>0)
        if (piece_colour(p_x-1,p_y)~= r_colour && chessboard(p_x-1,p_y)~=0)
            possiblemoves(p_x-1,p_y) = 2;
        elseif (piece_colour(p_x-1,p_y)== r_colour)
            ;
        else
            possiblemoves(p_x-1,p_y) = 1;
        end
    end

    if(p_x-1>0 && p_y+1<9)
        if (piece_colour(p_x-1,p_y+1)~= r_colour && chessboard(p_x-1,p_y+1)~=0)
            possiblemoves(p_x-1,p_y+1) = 2;
        elseif (piece_colour(p_x-1,p_y+1)== r_colour)
            ;
        else
            possiblemoves(p_x-1,p_y+1) = 1;
        end
    end

    if(p_x-1>0 && p_y-1>0)
        if (piece_colour(p_x-1,p_y-1)~= r_colour && chessboard(p_x-1,p_y-1)~=0)
            possiblemoves(p_x-1,p_y-1) = 2;
        elseif (piece_colour(p_x-1,p_y-1)== r_colour)
            ;
        else
            possiblemoves(p_x-1,p_y-1) = 1;
        end
    end

%-------------------------------------------------------------------------   
%                                  Castling 
%-------------------------------------------------------------------------

%---------------------------- For white king -----------------------------
%Checks to see if traversed squares are in check and if the final square is
%checked. Also checks if squares in between are empty
   if (piece_colour(p_x,p_y)==119 && num_moves(p_x,p_y)==0 && num_moves(8,8)==0 .... 
           && piece_colour(8,6)==0 && piece_colour(8,7)==0 && potential_moves(8,6)==0 ...  
           && potential_moves(8,7)==0 && chessboard(8,8)==5 && piece_colour(8,8)==119 )
       possiblemoves(8,7) = 4;  
   end    
   if (piece_colour(p_x,p_y)==119 && num_moves(p_x,p_y)==0 && num_moves(8,1)==0 ....
           && piece_colour(8,2)==0 && piece_colour(8,3)==0 && piece_colour(8,4)==0 ... 
           && potential_moves(8,3)==0 && potential_moves(8,4)==0 && chessboard(8,1)==5 ...
           && piece_colour(8,1)==119)
       possiblemoves(8,3) = 4;
   end

%---------------------------- For black king -----------------------------
%Checks to see if traversed squares are in check and if the final square is
%checked
   if (piece_colour(p_x,p_y)==98 && num_moves(p_x,p_y)==0 && num_moves(1,1)==0 .... 
           && piece_colour(1,2)==0 && piece_colour(1,3)==0 && piece_colour(1,4)==0 .... 
           && potential_moves(1,3)==0 && potential_moves(1,4)==0 && ...
           chessboard(1,1)==5 && piece_colour(1,1)==98)
       possiblemoves(1,3) = 4;
   end    
   if (piece_colour(p_x,p_y)==98 && num_moves(p_x,p_y)==0 && num_moves(1,8)==0 .... 
           && piece_colour(1,6)==0 && piece_colour(1,7)==0 && potential_moves(1,6)==0 .... 
           && potential_moves(1,7)==0 && chessboard(1,8)==5 && piece_colour(1,8)==98)
       possiblemoves(1,7) = 4;
   end

possiblemoves(p_x,p_y)=0;
%-------------------------------------------------------------------------
end
\end{lstlisting}

\subsubsection{MovementBishop.m}
\begin{lstlisting}
function [possiblemoves] = MovementBishop(chessboard,piece_colour,p_x,p_y)

%Initialisation values --------------------------------------------------
r_colour = piece_colour(p_x,p_y);
possiblemoves = zeros(8,8);

%This section allows movement in / direction -----------------------------
i=1;
while(p_x+i<9 && p_y+i<9)
    if(piece_colour(p_x+i,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y+i)~= r_colour && chessboard(p_x+i,p_y+i)~=0)
        possiblemoves(p_x+i,p_y+i) = 2;
        break
    end
    possiblemoves(p_x+i,p_y+i) = 1;
    i = i+1;
end

i=1;
while(p_x-i>0 && p_y-i>0)
    if(piece_colour(p_x-i,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x-i,p_y-i)~= r_colour && chessboard(p_x-i,p_y-i)~=0)
        possiblemoves(p_x-i,p_y-i) = 2;
        break
    end
    possiblemoves(p_x-i,p_y-i) = 1;
    i = i+1;
end

%This section allows movement in the \ direction-------------------------
i=1;
while(p_x+i<9 && p_y-i>0)
    if(piece_colour(p_x+i,p_y-i)== r_colour)
        break
    end
    if(piece_colour(p_x+i,p_y-i)~= r_colour && chessboard(p_x+i,p_y-i)~=0)
        possiblemoves(p_x+i,p_y-i) = 2;
        break
    end
    possiblemoves(p_x+i,p_y-i) = 1;
    i = i+1;
end

i=1;
while(p_x-i>0 && p_y+i<9)
    if(piece_colour(p_x-i,p_y+i)== r_colour)
        break
    end
    if(piece_colour(p_x-i,p_y+i)~= r_colour && chessboard(p_x-i,p_y+i)~=0)
        possiblemoves(p_x-i,p_y+i) = 2;
        break
    end
    possiblemoves(p_x-i,p_y+i) = 1;
    i = i+1;
end

%-------------------------------------------------------------------------

end
\end{lstlisting}

\subsection{Front-end}
\subsubsection{ClickPiece.m}
\begin{lstlisting}
%ClickPiece Obtains all the data from a user's click, highlights possible
%moves and allows the user to make that move.
function [varargout]=ClickPiece(var1,var2,B,piece_colour,chessboard,...
    num_moves,parameters,potentialmoves,handles,varargin )

set(handles.gameconsole,'String','')

%----------Determines which colour is able to be selected------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

onlyAIoption = 0;
%-------------------------------------------------------------------------
 clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph Grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%-------------------------------------------------------------------------
%This is the board
      piecetype = B.top(x,y).name;

%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4;
      p_y = y - 4;

if(piece_colour(p_x,p_y) == colourturn)
%----------------------Generates Possible Moves---------------------------

switch piecetype
    case 'pawn'
        [possiblemoves] = MovementPawn(chessboard,piece_colour,num_moves,p_x,p_y);
    case 'rook'
        [possiblemoves] = MovementRook(chessboard,piece_colour,p_x,p_y);
    case 'knight'
        [possiblemoves] = MovementKnight(chessboard,piece_colour,p_x,p_y);
    case 'bishop'
        [possiblemoves] = MovementBishop(chessboard,piece_colour,p_x,p_y);
    case 'queen'
        [possiblemoves] = MovementQueen(chessboard,piece_colour,p_x,p_y);
    case 'king'
        [possiblemoves] = MovementKing(chessboard,piece_colour,num_moves,...
            potentialmoves,p_x,p_y);
end

%-------------------------------------------------------------------------
%             REDRAWS THE BOARD BUT HIGHLIGHTS POSSIBLE MOVES
%-------------------------------------------------------------------------
%------------------------------Draws Rectangles---------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

%----------- Highlights possible moves------------------------------------
for r=1:parameters.rows
    for c=1:parameters.cols
        switch possiblemoves(r,c)
%_______________________Highlights movable squares________________________
            case 1
             rectangle('Position',[parameters.xx(9-r,c),parameters.yy(9-r,c),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],'FaceColor','y',...
                 'ButtonDownFcn',{@ClickMovePiece,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0})
%_______________________Highlights capturable squares______________________
            case 2
             rectangle('Position',[parameters.xx(9-r,c),parameters.yy(9-r,c),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],'FaceColor','r')
%_______________________Highlights Enpassant Squares_______________________
            case 3
             rectangle('Position',[parameters.xx(9-r,c),parameters.yy(9-r,c),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],'FaceColor','r',...
                 'ButtonDownFcn',{@ClickEnpassant,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0})
%_______________________Highlights Castling Squares________________________
            case 4
             rectangle('Position',[parameters.xx(9-r,c),parameters.yy(9-r,c),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],'FaceColor','b',...
                 'ButtonDownFcn',{@ClickCastling,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0})
%_______________________Highlights Pawn Promotion Square___________________
            case 5
             rectangle('Position',[parameters.xx(9-r,c),parameters.yy(9-r,c),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],'FaceColor','c',...
                 'ButtonDownFcn',{@ClickPawnPromo,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0,0})
         end
    end
end
%--------------------------------------------------------------------------
%                             Redraws images 
%--------------------------------------------------------------------------
for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            %If Statement enables capture move
            if possiblemoves(r,c) == 2
                imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickCapturePiece,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0});
             %Enables Pawn Promotion
            elseif possiblemoves(r,c) == 5 && chessboard(r,c)~=0
                imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPawnPromo,x,y,B,piece_colour,chessboard...
                 ,num_moves,parameters,possiblemoves,handles,onlyAIoption,0,0});
             %Else enable click piece
            else
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles,onlyAIoption,0,0});
            end
        end
    end
end
drawnow;
end
end
%-------------------------------------------------------------------------
%-------------------------------------------------------------------------

\end{lstlisting}

\subsubsection{ClickCapturePiece.m}
\begin{lstlisting}
%CapturePiece Part of the Click Series of Functions - Enables capture
function [chessboard,piece_colour, num_moves,allowscheck]=ClickCapturePiece(v1,v2,x_ori,y_ori,B,piece_colour,chessboard,...
    num_moves,parameters,PM,handles,onlyAIoption,move_x,move_y,varargin)

%--------------------------------------------------------------------------
%                  Init values,conversions and click location
%--------------------------------------------------------------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

if onlyAIoption == 0
clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4; %p_x is necessary because it is the current clicked position
      p_y = y - 4;
      ori_x = x_ori - 4; %The difference is that ori_x is for chessboard,
      ori_y = y_ori - 4; %x_ori is for B.top
else
    p_x = move_x;   %Where is it moving to
    p_y = move_y;
    ori_x = x_ori;   %Where was it originally
    ori_y = y_ori;
end

%-------------------------------------------------------------------------
%            Checks if King is exposed to check in any way
%-------------------------------------------------------------------------
%The method used is to create a future chessboard based on the move
%requested

fboard = chessboard;
f_p_colour= piece_colour;
f_num_moves = num_moves;
%This step officially moves the piece
fboard(p_x,p_y) = chessboard(ori_x,ori_y);
f_p_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
f_num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
%This step empties the previous box
fboard(ori_x,ori_y) = 0;
f_p_colour(ori_x,ori_y) = 0;
f_num_moves(ori_x,ori_y) = 0;

%Analyses the future board
[potentialfuturemoves,capt_index_future] = analyseboard(fboard,...
    f_p_colour,f_num_moves,oppositecolour);
[allowscheck]=KingCheck(fboard,f_p_colour,colourturn,...
    capt_index_future,potentialfuturemoves);
if allowscheck==1 && onlyAIoption == 0
    set(handles.gameconsole,'String','King will be left in check, move invalid')
end
%-------------------------------------------------------------------------
%Ensures it can only move legally
if PM(p_x,p_y)==2 && chessboard(p_x,p_y) ~= 10 && allowscheck==0

B.info.turn = B.info.turn + 1;
%-------------------------------------------------------------------------
%              This is to edit the backend chessboard matrix
%-------------------------------------------------------------------------
%This step officially moves the piece
chessboard = fboard;
piece_colour = f_p_colour; 
num_moves = f_num_moves;

%-------------To Check Opposing Side -------------------------------------
[potentialmoves,capt_index] = analyseboard(chessboard,piece_colour,num_moves,colourturn);
[checkopp]=KingCheck(chessboard,piece_colour,oppositecolour,capt_index,potentialmoves);
if checkopp == 1 && onlyAIoption == 0
   set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
elseif checkopp == 0 && onlyAIoption ==0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end

if onlyAIoption == 0
    [B] = readchessboard(B,chessboard,piece_colour);
%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles});
        end
    end
end
drawnow;
if(get(handles.choice2,'Value')==1)
    AIControl(B,piece_colour,chessboard,num_moves,parameters, handles);
end
if(get(handles.choice3,'Value')==1)
    AIvsAI(B,piece_colour,chessboard,num_moves,parameters, handles)
end
if(get(handles.choice1,'Value')==1)
    PlayerVsPlayer( B,piece_colour,chessboard,num_moves,parameters, handles )
end
end
%---------------------------------------------------------------------------------------
end
end
\end{lstlisting}

\subsubsection{ClickCastling.m}
\begin{lstlisting}
%Castling Enables frontend implementation of castling
function [chessboard,piece_colour, num_moves,allowscheck]=ClickCastling(v1,v2,x_ori,y_ori,B,piece_colour,chessboard,...
    num_moves,parameters,PM,handles,onlyAIoption,move_x,move_y,varargin)

%--------------------------------------------------------------------------
%                  Init values,conversions and click location
%--------------------------------------------------------------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

if onlyAIoption == 0
clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4; %p_x is necessary because it is the current clicked position
      p_y = y - 4;
      ori_x = x_ori - 4; %The difference is that ori_x is for chessboard,
      ori_y = y_ori - 4; %x_ori is for B.top
else
    p_x = move_x;   %Where is it moving to
    p_y = move_y;
    ori_x = x_ori;   %Where was it originally
    ori_y = y_ori;
end

%-------------------------------------------------------------------------
%            Checks if King is exposed to check in any way
%-------------------------------------------------------------------------
%The method used is to create a future chessboard based on the move
%requested

fboard = chessboard;
f_p_colour= piece_colour;
f_num_moves = num_moves;
%This step officially moves the piece
fboard(p_x,p_y) = chessboard(ori_x,ori_y);
f_p_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
f_num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
%This step empties the previous box
fboard(ori_x,ori_y) = 0;
f_p_colour(ori_x,ori_y) = 0;
f_num_moves(ori_x,ori_y) = 0;

%Analyses the future board
[potentialfuturemoves,capt_index_future] = analyseboard(fboard,...
    f_p_colour,f_num_moves,oppositecolour);
[allowscheck]=KingCheck(fboard,f_p_colour,colourturn,...
    capt_index_future,potentialfuturemoves);
if allowscheck ==1 && onlyAIoption == 0
    set(handles.gameconsole,'String','King will be left in check, move invalid')
end
%-------------------------------------------------------------------------
%Ensures it can only move legally
if PM(p_x,p_y)==4 && allowscheck==0

%-------------------------------------------------------------------------
%                   B.top
%-------------------------------------------------------------------------
%Coordinate system is X_rook = [B.top Chessboard]
if ( p_x == 8 && p_y == 7)
    x_rook = [12 8]; %Initial Rook Position
    y_rook = [12 8];
    move_x = [12 8]; %Final Rook Position
    move_y = [10 6];
elseif ( p_x == 8 && p_y == 3 )
    x_rook = [12 8];
    y_rook = [5 1];
    move_x = [12 8];
    move_y = [8 4];
elseif ( p_x == 1 && p_y == 7)
    x_rook = [5 1];
    y_rook = [12 8];
    move_x = [5 1];
    move_y = [10 6];
elseif ( p_x == 1 && p_y == 3)
    x_rook = [5 1];
    y_rook = [5 1];
    move_x = [5 1];
    move_y = [8 4];
end

B.info.turn = B.info.turn + 1;
%-------------------------------------------------------------------------
%              This is to edit the backend chessboard matrix
%-------------------------------------------------------------------------
%--------------------------------King-------------------------------------
%This step officially moves the piece
chessboard(p_x,p_y) = chessboard(ori_x,ori_y);
piece_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;

%This step empties the previous box
chessboard(ori_x,ori_y) = 0;
piece_colour(ori_x,ori_y) = 0;
num_moves(ori_x,ori_y) = 0;

%-------------------------------Rook--------------------------------------
%This step officially moves the piece
chessboard(move_x(2),move_y(2)) = chessboard(x_rook(2),y_rook(2));
piece_colour(move_x(2),move_y(2)) = piece_colour(x_rook(2),y_rook(2));
num_moves(move_x(2),move_y(2)) = num_moves(x_rook(2),y_rook(2)) + 1;

%This step empties the previous box
chessboard(x_rook(2),y_rook(2)) = 0;
piece_colour(x_rook(2),y_rook(2)) = 0;
num_moves(x_rook(2),y_rook(2)) = 0;

%-------------Analyses for potential checks & provides game stats---------
[potentialmoves,capt_index] = analyseboard(chessboard,piece_colour,num_moves,colourturn);
[checkopp]=KingCheck(chessboard,piece_colour,oppositecolour,capt_index,potentialmoves);
if checkopp == 1 && onlyAIoption == 0
    set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
    elseif checkopp == 0 && onlyAIoption ==0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end

if onlyAIoption ==0
    [B] = readchessboard(B,chessboard,piece_colour);
%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles});
        end
    end
end
drawnow;
if(get(handles.choice2,'Value')==1)
    AIControl(B,piece_colour,chessboard,num_moves,parameters, handles);
end
if(get(handles.choice3,'Value')==1)
    AIvsAI(B,piece_colour,chessboard,num_moves,parameters, handles)
end
if(get(handles.choice1,'Value')==1)
    PlayerVsPlayer( B,piece_colour,chessboard,num_moves,parameters, handles )
end
end
%---------------------------------------------------------------------------------------
end
end
\end{lstlisting}

\subsubsection{ClickEnpassant.m}
\begin{lstlisting}
%Enpassant Enables frontend implementation of En Passant
function [chessboard,piece_colour, num_moves,allowscheck]=ClickEnpassant(v1,v2,x_ori,y_ori,B,piece_colour,chessboard,...
    num_moves,parameters,PM, handles,onlyAIoption,move_x,move_y,varargin)

%--------------------------------------------------------------------------
%                  Init values,conversions and click location
%--------------------------------------------------------------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

if onlyAIoption == 0
clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4; %p_x is necessary because it is the current clicked position
      p_y = y - 4;
      ori_x = x_ori - 4; %The difference is that ori_x is for chessboard,
      ori_y = y_ori - 4; %x_ori is for B.top
else
    p_x = move_x;   %Where is it moving to
    p_y = move_y;
    ori_x = x_ori;   %Where was it originally
    ori_y = y_ori;
end

%-------------------------------------------------------------------------
%        Checks if King is exposed to check in any way due to move
%-------------------------------------------------------------------------
%The method used is to create a future chessboard based on the move
%requested

fboard = chessboard;
f_p_colour= piece_colour;
f_num_moves = num_moves;
%This step officially moves the piece
fboard(p_x,p_y) = chessboard(ori_x,ori_y);
f_p_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
f_num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
%This step empties the previous box
fboard(ori_x,ori_y) = 0;
f_p_colour(ori_x,ori_y) = 0;
f_num_moves(ori_x,ori_y) = 0;

%Analyses the future board
[potentialfuturemoves,capt_index_future] = analyseboard(fboard,...
    f_p_colour,f_num_moves,oppositecolour);
[allowscheck]=KingCheck(fboard,f_p_colour,colourturn,...
    capt_index_future,potentialfuturemoves);
if allowscheck==1 && onlyAIoption == 0
    set(handles.gameconsole,'String','King will be left in check, move invalid')
end
%-------------------------------------------------------------------------
%Ensures it can only move legally
if PM(p_x,p_y)==3 && allowscheck==0
%--------------------------------------------------------------------------
%                Moves Data in B.TOP & deletes previous cell
%--------------------------------------------------------------------------
%Coordinates of the captured piece
if (piece_colour(ori_x,ori_y)==98)
   del_x = [p_x+3 p_x-1];
   del_y = [p_y+4 p_y];
end

if (piece_colour(ori_x,ori_y)==119)    
   del_x = [p_x+5 p_x+1];
   del_y = [p_y+4 p_y];
end

B.info.turn = B.info.turn + 1;
%-------------------------------------------------------------------------
%              This is to edit the backend chessboard matrix
%-------------------------------------------------------------------------
%This step officially moves the piece
chessboard(p_x,p_y) = chessboard(ori_x,ori_y);
piece_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;

%This step empties the previous box
chessboard(ori_x,ori_y) = 0;
piece_colour(ori_x,ori_y) = 0;
num_moves(ori_x,ori_y) = 0;

%This step deletes the capured piece
chessboard(del_x(2),del_y(2)) = 0;
piece_colour(del_x(2),del_y(2)) = 0;
num_moves(del_x(2),del_y(2)) = 0;

%-------------Analyses for potential checks & provides game stats---------
[potentialmoves,capt_index] = analyseboard(chessboard,piece_colour,num_moves,colourturn);
[checkopp]=KingCheck(chessboard,piece_colour,oppositecolour,capt_index,potentialmoves);
if checkopp == 1 && onlyAIoption == 0
    set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
    elseif checkopp == 0 && onlyAIoption ==0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end

if onlyAIoption == 0
    [B] = readchessboard(B,chessboard,piece_colour);
%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles});
        end
    end
end
drawnow;
if(get(handles.choice2,'Value')==1)
    AIControl(B,piece_colour,chessboard,num_moves,parameters, handles);
end
if(get(handles.choice3,'Value')==1)
    AIvsAI(B,piece_colour,chessboard,num_moves,parameters, handles)
end
if(get(handles.choice1,'Value')==1)
    PlayerVsPlayer( B,piece_colour,chessboard,num_moves,parameters, handles )
end
end
%---------------------------------------------------------------------------------------
end
end
\end{lstlisting}

\subsubsection{ClickPawnPromo.m}
\begin{lstlisting}
%PawnPromo Enables Front End Implementation of Pawn Promo
function [chessboard,piece_colour, num_moves,allowscheck]=ClickPawnPromo(v1,v2,x_ori,y_ori,B,piece_colour,chessboard,...
    num_moves,parameters,PM,handles,onlyAIoption,move_x,move_y,promo,varargin)

%--------------------------------------------------------------------------
%                  Init values,conversions and click location
%--------------------------------------------------------------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

if onlyAIoption == 0
clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4; %p_x is necessary because it is the current clicked position
      p_y = y - 4;
      ori_x = x_ori - 4; %The difference is that ori_x is for chessboard,
      ori_y = y_ori - 4; %x_ori is for B.top
else
    p_x = move_x;   %Where is it moving to
    p_y = move_y;
    ori_x = x_ori;   %Where was it originally
    ori_y = y_ori;
end

%-------------------------------------------------------------------------
%            Checks if King is exposed to check in any way
%-------------------------------------------------------------------------
%The method used is to create a future chessboard based on the move
%requested

fboard = chessboard;
f_p_colour= piece_colour;
f_num_moves = num_moves;
%This step officially moves the piece
fboard(p_x,p_y) = chessboard(ori_x,ori_y);
f_p_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
f_num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
%This step empties the previous box
fboard(ori_x,ori_y) = 0;
f_p_colour(ori_x,ori_y) = 0;
f_num_moves(ori_x,ori_y) = 0;

%Analyses the future board
[potentialfuturemoves,capt_index_future] = analyseboard(fboard,...
    f_p_colour,f_num_moves,oppositecolour);
[allowscheck]=KingCheck(fboard,f_p_colour,colourturn,...
    capt_index_future,potentialfuturemoves);
if allowscheck==1 && onlyAIoption == 0
    set(handles.gameconsole,'String','King will be left in check, move invalid')
end
%-------------------------------------------------------------------------
%Ensures it can only move legally
if PM(p_x,p_y)==5 && allowscheck==0
%--------------------------------------------------------------------------
%                Moves Data in B.TOP & deletes previous cell
%--------------------------------------------------------------------------
%Allows user to input desired piece. Checks legality.
if ~onlyAIoption
set(handles.gameconsole,'String','Pawn has been promoted');
            flags=0;
          while(flags==0)
              flags=1;
              v=0;
              while v == 0
              [pawn_prom,v] = listdlg('PromptString','Select a piece:',...
                'SelectionMode','single',...
                'ListString',{'Rook','Queen','Knight','Bishop'});
              end
            switch pawn_prom
                case 1
                    chessboard(p_x,p_y)= 5;
                case 2
                    chessboard(p_x,p_y)= 9;
                case 3
                    chessboard(p_x,p_y)= 3;
                case 4
                    chessboard(p_x,p_y)= 4;
                otherwise
                    disp('Invalid input');
                    flags=0;
            end   
          end
else
            switch promo
                        case 'rook'
                            chessboard(p_x,p_y)= 5;
                        case 'queen'
                            chessboard(p_x,p_y)= 9;
                        case 'knight'
                            chessboard(p_x,p_y)= 3;
                        case 'bishop'
                            chessboard(p_x,p_y)= 4;
            end   
end

B.info.turn = B.info.turn + 1;

%-------------------------------------------------------------------------
%              This is to edit the backend chessboard matrix
%-------------------------------------------------------------------------
%This step officially moves the piece
num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
piece_colour(p_x,p_y)= colourturn;

%This step empties the previous box
chessboard(ori_x,ori_y) = 0;
piece_colour(ori_x,ori_y) = 0;
num_moves(ori_x,ori_y) = 0;

%-------------Analyses for potential checks & provides game stats---------
[potentialmoves,capt_index] = analyseboard(chessboard,piece_colour,num_moves,colourturn);
[checkopp]=KingCheck(chessboard,piece_colour,oppositecolour,capt_index,potentialmoves);
if checkopp == 1 && onlyAIoption == 0
   set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
elseif checkopp == 0 && onlyAIoption ==0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end

if onlyAIoption == 0
    [B] = readchessboard(B,chessboard,piece_colour);
%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles});
        end
    end
end
drawnow;
if(get(handles.choice2,'Value')==1)
    AIControl(B,piece_colour,chessboard,num_moves,parameters, handles);
end
if(get(handles.choice3,'Value')==1)
    AIvsAI(B,piece_colour,chessboard,num_moves,parameters, handles)
end
if(get(handles.choice1,'Value')==1)
    PlayerVsPlayer( B,piece_colour,chessboard,num_moves,parameters, handles )
end
end
%---------------------------------------------------------------------------------------
end
end
\end{lstlisting}

\subsubsection{ClickMovePiece.m}
\begin{lstlisting}
%Movepiece Part of the Click Series of Functions - Enables movement
function [chessboard,piece_colour, num_moves,allowscheck]=ClickMovePiece(v1,v2,x_ori,y_ori,B,piece_colour,chessboard,...
    num_moves,parameters,PM,handles,onlyAIoption,move_x,move_y,varargin)

%--------------------------------------------------------------------------
%                  Init values,conversions and click location
%--------------------------------------------------------------------------
if(mod(B.info.turn,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

if onlyAIoption == 0
clickP = get(gca,'CurrentPoint');
      x = ceil(clickP(1,2));
      y = ceil(clickP(1,1));
%---- Conversion from Graph grid to B.top grid ---------------------------
      x = 13-x;      
      y = y + 4;
%--------Conversion from B.Top grid to Chessboard grid--------------------
      p_x = x - 4; %p_x is necessary because it is the current clicked position
      p_y = y - 4;
      ori_x = x_ori - 4; %The difference is that ori_x is for chessboard,
      ori_y = y_ori - 4; %x_ori is for B.top
else
    p_x = move_x;   %Where is it moving to
    p_y = move_y;
    ori_x = x_ori;   %Where was it originally
    ori_y = y_ori;
end

%-------------------------------------------------------------------------
%            Checks if King is exposed to check in any way
%-------------------------------------------------------------------------
%The method used is to create a future chessboard based on the move
%requested

fboard = chessboard;
f_p_colour= piece_colour;
f_num_moves = num_moves;
%This step officially moves the piece
fboard(p_x,p_y) = chessboard(ori_x,ori_y);
f_p_colour(p_x,p_y) = piece_colour(ori_x,ori_y);
f_num_moves(p_x,p_y) = num_moves(ori_x,ori_y) + 1;
%This step empties the previous box
fboard(ori_x,ori_y) = 0;
f_p_colour(ori_x,ori_y) = 0;
f_num_moves(ori_x,ori_y) = 0;

%Analyses the future board
[potentialfuturemoves,capt_index_future] = analyseboard(fboard,...
    f_p_colour,f_num_moves,oppositecolour);
[allowscheck]=KingCheck(fboard,f_p_colour,colourturn,...
    capt_index_future,potentialfuturemoves);
if allowscheck ==1 && onlyAIoption == 0
    set(handles.gameconsole,'String','King will be left in check, move invalid')
end
%-------------------------------------------------------------------------
%------------------------------------------------------------------------

if PM(p_x,p_y)==1 && allowscheck==0 %Ensures it can only move legally

%Iterates the turn
B.info.turn = B.info.turn + 1;

%-------------------------------------------------------------------------
%              This is to edit the backend chessboard matrix
%-------------------------------------------------------------------------
%This step officially moves the piece
chessboard = fboard;
piece_colour = f_p_colour; 
num_moves = f_num_moves;

%-------------Analyses for potential checks & provides game stats---------
[potentialmoves,capt_index] = analyseboard(chessboard,piece_colour,num_moves,colourturn);
[checkopp]=KingCheck(chessboard,piece_colour,oppositecolour,capt_index,potentialmoves);

if checkopp == 1 && onlyAIoption ==0
    set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
elseif checkopp == 0 && onlyAIoption ==0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end

%-------------------------------------------------------------------------
if onlyAIoption == 0
    [B] = readchessboard(B,chessboard,piece_colour);
%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

%------------------------------------------------------------------------
for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,potentialmoves,handles});
        end
    end
end
drawnow;
if(get(handles.choice2,'Value')==1)
    AIControl(B,piece_colour,chessboard,num_moves,parameters, handles);
end
if(get(handles.choice3,'Value')==1)
    AIvsAI(B,piece_colour,chessboard,num_moves,parameters, handles)
end
if(get(handles.choice1,'Value')==1)
    PlayerVsPlayer( B,piece_colour,chessboard,num_moves,parameters, handles )
end
end
%-------------------------------------------------------------------------
end
end
\end{lstlisting}

\subsection{AI}
\subsubsection{AIControl.m}
\begin{lstlisting}
function [B,piece_colour,chessboard,num_moves,parameters, handles]=AIControl(B,piece_colour,chessboard,...
                num_moves,parameters, handles)
%AIControl Enables AI to be in action

%-------------------------------------------------------------------------
%                       Init Values
%-------------------------------------------------------------------------
if(mod(B.info.turn-1,2)==1)
    colourturn = 119;
    oppositecolour = 98;
else
    colourturn = 98;
    oppositecolour = 119;
end

[userboardscore] = heuristicanalysis(B,chessboard, piece_colour,num_moves,119,handles);
set(handles.UPS,'String',userboardscore)
handles.userboardscore = [handles.userboardscore userboardscore];
depth = 2;
set(handles.depth,'String',depth)

%------------------ Stops Game Execution if White Wins -------------------
% [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
% if ischeckmate
%     return
% end

[oppcolourpotentialmoves,oppcolourcapt_index] = analyseboard(chessboard, piece_colour,num_moves,colourturn);

[ischeck]=KingCheck(chessboard,piece_colour,oppositecolour, oppcolourcapt_index,oppcolourpotentialmoves);
if ischeck == 1
    set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, White Wins')
    end
elseif ischeck == 0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end
%--------------------Plot UserBoardScore-----------------------------------

 handles.turnforwhite = [handles.turnforwhite B.info.turn];
 plot(handles.graph,handles.turnforwhite,handles.userboardscore,'-b',...
     handles.turnforblack,handles.AIBoardscore,'-r','LineWidth',2)
 set(handles.graph,'XColor','w','YColor','w')
 xlabel(handles.graph,'Turn')
 ylabel(handles.graph,'Score')

%-------------------------------------------------------------------------
set(handles.AIMsgs,'String','Thinking Really Hard')

%Produces AI's decision
tic
[boardscore,chessboard,piece_colour,num_moves]=...
    AI_GenerateAllMoves(B,chessboard,piece_colour,num_moves,depth,1,-99999,99999,handles);
time =toc;

set(handles.AIMsgs,'String',['Time Taken To Think Was: ' num2str(time) ' seconds'])

%Translates the results into B.top
[B] = readchessboard(B,chessboard,piece_colour);
%Iterates turn
B.info.turn = B.info.turn + 1;

%--------------------- Shows AI Board Score-------------------------------
[AIBoardScore] = heuristicanalysis(B,chessboard, piece_colour,num_moves,98,handles);
set(handles.APS,'String',AIBoardScore)
handles.AIBoardscore = [handles.AIBoardscore AIBoardScore];

%---------------------Plots AI Board Score--------------------------------

 handles.turnforblack = [handles.turnforblack B.info.turn];
 plot(handles.graph,handles.turnforwhite,handles.userboardscore,'-b',...
     handles.turnforblack,handles.AIBoardscore,'-r','LineWidth',2)
 set(handles.graph,'XColor','w','YColor','w')
 xlabel(handles.graph,'Turn')
 ylabel(handles.graph,'Score')

%------------------------ Checks if AI has checkmated User ---------------
[oppcolourpotentialmoves,oppcolourcapt_index] = analyseboard(chessboard, piece_colour,num_moves,oppositecolour);

[ischeck]=KingCheck(chessboard,piece_colour,colourturn, oppcolourcapt_index,oppcolourpotentialmoves);
if ischeck == 1
    set(handles.checkstat,'String','Check')
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Checkmate, Black Wins')
    end
elseif ischeck == 0
    [ischeckmate]=checkmate(B,chessboard,piece_colour, num_moves);
    if ischeckmate
        set(handles.checkstat,'String','Stalemate')
    else
        set(handles.checkstat,'String','')
    end
end
%-------------------------------------------------------------------------

%-------------------------------------------------------------------------
%                           Redraws the Board
%-------------------------------------------------------------------------
icount=0;
for i=1:71
         icount=icount+1;
         if mod(i,2)==1
             rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                 parameters.dx ,parameters.dx],'Curvature',[0,0],...
                 'FaceColor',[0.82 0.545 0.278])
         else
            rectangle('Position',[parameters.xx(icount),parameters.yy(icount),...
                parameters.dx ,parameters.dx],...
                'Curvature',[0,0],'FaceColor',[1 0.808 0.62])             
         end
end

%------------------------------------------------------------------------
for r=1:parameters.rows
    for c=1:parameters.cols
        if ~isempty(B.top(r+B.info.pad/2,c+B.info.pad/2).image)
            % load the image
            [X, map, alpha]  = imread(B.top(r+B.info.pad/2,c+B.info.pad/2).image);
            % draw the image
            imHdls(r,c) = image(c+[0 1]-1,[parameters.rows-1 parameters.rows]-r+1,...
                mirrorImage(X),'AlphaData',mirrorImage(alpha),...
                'ButtonDownFcn',{@ClickPiece,B,piece_colour,chessboard,...
                num_moves,parameters,oppcolourpotentialmoves,handles});
        end
    end
end
drawnow;
%-------------------------------------------------------------------------
end
\end{lstlisting}

\subsubsection{AI\_GenerateAllMoves.m}
\begin{lstlisting}
%AI - Generates moves and stores them for 1 PLY (Only for DATA Tree)
function [boardscore,bchessboard,bpiece_colour,bnum_moves,handles]=...
    AI_GenerateAllMoves(B,chessboard,piece_colour,num_moves,depth,maxormin,alpha,beta,handles)
%-------------------------------------------------------------------------
%                           Init Values
%-------------------------------------------------------------------------
TmpB = B;

if(mod(TmpB.info.turn,2)==1)
    colour = 119;
    oppcolour = 98;
else
    colour = 98;
    oppcolour = 119;
end

TmpB.info.turn = TmpB.info.turn +1;
%-------------------------------------------------------------------------

if depth == 0
   TmpB.info.turn = TmpB.info.turn-1;
  [boardscore] = heuristicanalysis(TmpB,chessboard, piece_colour,num_moves,colour,handles);
  bchessboard = chessboard;
  bpiece_colour = piece_colour;
  bnum_moves = num_moves;
else

if maxormin == 1  %Maximizing Player
%===================== Generates Future Nodes or Leafs ====================
%-------------------------------------------------------------------------
%        Loop that generates all possible moves
%-------------------------------------------------------------------------
[p_x,p_y] = find(piece_colour == colour);
perm_index = randperm(length(p_x));
p_x = p_x(perm_index);
p_y = p_y(perm_index);
n_remaining = length(p_x);
[potentialmoves] = analyseboard(chessboard, piece_colour,num_moves,oppcolour);
previousboardscore = -99999;
%In essence, we are going through each piece, looking at it's possible
%moves, make those possible moves, evaluate, save bestboard.
for i=1:n_remaining
    p_type = chessboard(p_x(i),p_y(i));
    switch p_type
        case 1
            [move] = MovementPawn(chessboard,piece_colour,num_moves,p_x(i),p_y(i));
        case 5
            [move] = MovementRook(chessboard,piece_colour,p_x(i),p_y(i)); 
        case 4
            [move] = MovementBishop(chessboard,piece_colour,p_x(i),p_y(i));
        case 3
            [move] = MovementKnight(chessboard,piece_colour,p_x(i),p_y(i));
        case 9
            [move] = MovementQueen(chessboard,piece_colour,p_x(i),p_y(i));
        case 10
            [move] = MovementKing(chessboard,piece_colour,num_moves,potentialmoves,p_x(i),p_y(i));
    end

%-------------------------------------------------------------------------
%            Individual Piece Moves That Generate New Game States
%                    Recursion is also added in each loop
%-------------------------------------------------------------------------
     [move_x,move_y] = find(move ~= 0);
     perm_index2 = randperm(length(move_x));
     move_x = move_x(perm_index2);
     move_y = move_y(perm_index2);
     n_move = length(move_x);
     pruneflag = 0;
%This loop generates all the game states from 1 piece
     for j = 1:n_move
         switch move(move_x(j),move_y(j))
             case 1
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickMovePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 2
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCapturePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 3
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickEnpassant(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move, 0,1,move_x(j),move_y(j));
             case 4
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCastling(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 5
                 [pchessboard,ppiece_colour, pnum_moves,kingincheck]=ClickPawnPromo(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j),'queen');
         end
%--------A node has been generated, what do you want to do with it?--------
         if kingincheck  
             %ignore because move not valid
             if ~exist('boardscore','var')
             boardscore = -99999;
             bchessboard = 0;
             bpiece_colour =0;
             bnum_moves =0;
             end
         else
                 %Generate another layer with recursive parameters
                 [boardscore,~,~,~,handles]=...
            AI_GenerateAllMoves(TmpB,pchessboard,ppiece_colour,pnum_moves,depth-1,-maxormin,alpha,beta,handles);

                    if boardscore > previousboardscore
                        previousboardscore = boardscore;
                        bchessboard = pchessboard;
                        bpiece_colour = ppiece_colour;
                        bnum_moves = pnum_moves;
                    end
                    if boardscore>alpha
                        alpha = boardscore;
                    end
%             disp([depth alpha beta boardscore previousboardscore i j n_remaining n_move])
                    if alpha>beta
                        pruneflag = 1;
                        break
                    end
         end
%--------------------------------------------------------------------------
     end
     if pruneflag
         break
     end
end
%=========================================================================

elseif maxormin == -1 %Minimizing Player
%===================== Generates Future Nodes or Leafs ====================
%-------------------------------------------------------------------------
%        Loop that generates all possible moves
%-------------------------------------------------------------------------
[p_x,p_y] = find(piece_colour == colour);
perm_index = randperm(length(p_x));
p_x = p_x(perm_index);
p_y = p_y(perm_index);
n_remaining = length(p_x);
[potentialmoves] = analyseboard(chessboard, piece_colour,num_moves,oppcolour);
previousboardscore = 99999;
%In essence, we are going through each piece, looking at it's possible
%moves, make those possible moves, evaluate, save bestboard.
for i=1:n_remaining
    p_type = chessboard(p_x(i),p_y(i));
    switch p_type
        case 1
            [move] = MovementPawn(chessboard,piece_colour,num_moves,p_x(i),p_y(i));
        case 5
            [move] = MovementRook(chessboard,piece_colour,p_x(i),p_y(i)); 
        case 4
            [move] = MovementBishop(chessboard,piece_colour,p_x(i),p_y(i));
        case 3
            [move] = MovementKnight(chessboard,piece_colour,p_x(i),p_y(i));
        case 9
            [move] = MovementQueen(chessboard,piece_colour,p_x(i),p_y(i));
        case 10
            [move] = MovementKing(chessboard,piece_colour,num_moves,potentialmoves,p_x(i),p_y(i));
    end

%-------------------------------------------------------------------------
%            Individual Piece Moves That Generate New Game States
%                    Recursion is also added in each loop
%-------------------------------------------------------------------------
     [move_x,move_y] = find(move ~= 0);
     perm_index2 = randperm(length(move_x));
     move_x = move_x(perm_index2);
     move_y = move_y(perm_index2);
     n_move = length(move_x);
     pruneflag = 0;
%This loop generates all the game states from 1 piece
     for j = 1:n_move
         switch move(move_x(j),move_y(j))
             case 1
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickMovePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 2
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCapturePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 3
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickEnpassant(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move, 0,1,move_x(j),move_y(j));
             case 4
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCastling(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 5
                 [pchessboard,ppiece_colour, pnum_moves,kingincheck]=ClickPawnPromo(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j),'queen');
         end
%--------A node has been generated, what do you want to do with it?--------
         if kingincheck  
             %ignore because move not valid
             if ~exist('boardscore','var')
             boardscore = 99999;
             bchessboard = 0;
             bpiece_colour =0;
             bnum_moves =0;
             end
         else
                 %Generate another layer with recursive parameters
                 [boardscore,~,~,~,handles]=...
            AI_GenerateAllMoves(TmpB,pchessboard,ppiece_colour,pnum_moves,depth-1,-maxormin,alpha,beta,handles);

                if boardscore < previousboardscore
                    previousboardscore = boardscore;
                    bchessboard = pchessboard;
                    bpiece_colour = ppiece_colour;
                    bnum_moves = pnum_moves;
                end
                if boardscore<beta
                    beta = boardscore;
                end

% disp([depth alpha beta boardscore previousboardscore i j n_remaining n_move])
                if alpha>beta
                    pruneflag = 1;
                    break
                end
         end
     end
      if pruneflag
         break
     end
%--------------------------------------------------------------------------
end
%=========================================================================
end % For if maxormin
end % For DEPTH if condition
end %For Function
\end{lstlisting}

\subsubsection{heuristicanalysis.m}
\begin{lstlisting}
function [boardscore] = heuristicanalysis(B,chessboard, piece_colour,num_moves,currentcolour,handles)
%Colour should be the side in which it is being analysed for

%-------------------------------------------------------------------------
%                        Init Values
%-------------------------------------------------------------------------
if currentcolour == 119
    oppcolour = 98;
else
    oppcolour = 119;
end
%Generates potential moves of the currently investigated game state colour
[potentialmoves,capt_index] = analyseboard(chessboard, piece_colour,num_moves,currentcolour);
%Generates potential moves of the opponent
[oppcolourpotentialmoves, oppcolourcapt_index] = analyseboard(chessboard, piece_colour,num_moves,oppcolour);

%Finds the locations of own pieces and opponent's pieces
piece_index = find(piece_colour==currentcolour);
opp_piece_index = find(piece_colour==oppcolour);

%-------------------------------------------------------------------------

%-------------------Capture Analysis--------------------------------------
%A move is good because it opens up capture possibilities
num_pot_capture = length(capt_index); %Number of potential Captures
capt_value_sum = sum(chessboard(capt_index)); %The total capture value

%A move is good if it increases the number of capture 
capt_value_diff = 51 - sum(chessboard(opp_piece_index));

%------------------ Moves Analysis ---------------------------------------
%A move is good because it opens up space for other pieces to move
nocapture = potentialmoves;
nocapture(capt_index) = 0;
num_moves_available = sum(sum(nocapture));

%------------------ Threats ----------------------------------------------
%If the move causes other pieces to be under threat, the move is worse.
opp_num_pot_capture = length(oppcolourcapt_index);
opp_capt_value_sum = sum(chessboard(oppcolourcapt_index));

%------------------ Number of own pieces ---------------------------------
%A move is good if it prevents the number of own pieces from decreasing.
own_piece_sum_diff = 51 - sum(chessboard(piece_index));

%------------------- Control of centre space -----------------------------
%A move is good if it increases control of the centre of the board
centre_piece=zeros(8,8);
centre_piece([28 29 36 37])=chessboard([28 29 36 37]);
centre_piece = centre_piece~=0;
centre_space_sum =  sum(centre_piece(piece_index));

%------------------- Own King Checked? -----------------------------------
%Checks if own king is in check. If in check, also checks if its a checkmate
own_ischeck = KingCheck(chessboard,piece_colour,currentcolour,oppcolourcapt_index,oppcolourpotentialmoves);
if own_ischeck==1
    own_ischeckmate = checkmate(B,chessboard,piece_colour,num_moves);
    else own_ischeckmate = 0;
end

%--------------------------- Castling? -----------------------------------
%Checks if castling has taken place
rook_pos = find(chessboard==5 & piece_colour==currentcolour);
king_pos = find(chessboard==10 & piece_colour==currentcolour);
castle = 0;

if currentcolour == 98 %Black case
    if (king_pos==49 && ismember(41,rook_pos) && num_moves(41)==1 && num_moves(49)==1)
        castle = 1;
    elseif (king_pos==17 && ismember(25,rook_pos) && num_moves(25)==1 && num_moves(17)==1)
            castle = 1;
    end

else %White case
    if (king_pos==56 && ismember(48,rook_pos) && num_moves(48)==1 && num_moves(56)==1)
        castle = 1;
    elseif (king_pos==24 && ismember(32,rook_pos) && num_moves(32)==1 && num_moves(24)==1)
            castle = 1;
    end
end

%--------------------- Opponent Checkmate?--------------------------------
%Checks if opponent king is in check. If in check, also checks if its a checkmate
opp_ischeck = KingCheck(chessboard,piece_colour,oppcolour,capt_index,potentialmoves);
if opp_ischeck==1
    opp_ischeckmate = checkmate(B,chessboard,piece_colour,num_moves);
else opp_ischeckmate = 0;
end

%------------------- Possibility of opponenet's promotion? ---------------
%A move is bad if it brings opponent's pawn closer to the end of the board for promotion.
pawn_index = find(chessboard==1 & piece_colour==oppcolour);
if oppcolour == 98 %Black case
    end_dist = 8-rem(pawn_index,8);
    sum_opp_pawn_dist = sum(end_dist==0) + 0.5*sum(end_dist==1);
else %White case
    end_dist = rem(pawn_index,8)-1; 
    sum_opp_pawn_dist = sum(end_dist==0) + 0.5*sum(end_dist==1);
end

%--------------------- Possibility of own promotion? ---------------------
%A move is good if it brings own pawn closer to the end of the board for promotion.
pawn_index = find(chessboard==1 & piece_colour==currentcolour);
if currentcolour == 98 %Black case
    end_dist = 8-rem(pawn_index,8);
    sum_own_pawn_dist = sum(end_dist==0) + 0.5*sum(end_dist==1);
else %White case
    end_dist = rem(pawn_index,8)-1; 
    sum_own_pawn_dist = sum(end_dist==0) + 0.5*sum(end_dist==1);
end

%------------------ Gain Factor for Hard-------------------------------------
if(get(handles.setHard,'Value')==1)
gainCapture = 3;  %Encourages AI to position a piece such that it can capture more pieces in the next move
gainMoves = 10; %Encourages AI to position such that it opens space for other pieces
gainThreats = -4; %Discourages AI to make moves that will lead to threats
gainOpppieces = 25; %Encourages to make moves that decrease opponents pieces
gainOwnpieces = -5; %Discourages AI from making moves that decrease own pieces
gainCentre = 1; %Encourages AI to increase control of centre space
gainOwnprom = 1; %Encourages AI to promote own pawns close to the end of the board
gainOppprom = -10; %Discourages AI to promote opponent's pawns
end
%--------------------Gain Factor for Easy ---------------------------------
if(get(handles.setEasy,'Value')==1)
gainCapture = 2;  %Encourages AI to position a piece such that it can capture more pieces in the next move
gainMoves = 10; %Encourages AI to position such that it opens space for other pieces
gainThreats = -2; %Discourages AI to make moves that will lead to threats
gainOpppieces = 3.5; %Encourages to make moves that decrease opponents pieces
gainOwnpieces = 1; %Discourages AI from making moves that decrease own pieces
gainCentre = 10; %Encourages AI to increase control of centre space
gainOwnprom = 10; %Encourages AI to promote own pawns close to the end of the board
gainOppprom = -1; %Discourages AI to promote opponent's pawns
end
%--------------------- Gain Factor for Random -----------------------------
if(get(handles.setRandom,'Value')==1)
gainCapture = 0;  %Encourages AI to position a piece such that it can capture more pieces in the next move
gainMoves = 0; %Encourages AI to position such that it opens space for other pieces
gainThreats = 0; %Discourages AI to make moves that will lead to threats
gainOpppieces = 0; %Encourages to make moves that decrease opponents pieces
gainOwnpieces = 0; %Discourages AI from making moves that decrease own pieces
gainCentre = 0; %Encourages AI to increase control of centre space
gainOwnprom = 0; %Encourages AI to promote own pawns close to the end of the board
gainOppprom = 0; %Discourages AI to promote opponent's pawns
end
%----------------- Final Score Calculation ------------------------------
boardscore =  gainCapture * capt_value_sum... 
         + gainMoves * num_moves_available... 
         + gainThreats * opp_capt_value_sum...
         + gainOpppieces * capt_value_diff...
         + gainOwnpieces * own_piece_sum_diff...
         + gainCentre * centre_space_sum...
         + gainOwnprom * sum_own_pawn_dist...
         + gainOppprom * sum_opp_pawn_dist;
%Checks if castling has occured
    if castle == 1
        boardscore = boardscore + 250;
    end
%If a checkmate has occured, new boadscores are assigned
if(get(handles.setHard,'Value')==1 || get(handles.setEasy,'Value')==1 )
    if opp_ischeckmate == 1
        boardscore = 99999;
    end

    if own_ischeckmate == 1
        boardscore = -99999;
    end
end

if(get(handles.setRandom,'Value')==1)
    boardscore=rand* 2000;
end
end
\end{lstlisting}

\subsection{Board analysis}
\subsubsection{analyseboard.m}
\begin{lstlisting}
%Analyseboard Looks at one colour, sees where each piece is able to
%move. This is to allow for the Check function and castling.
%Colour in this case can be either current one or opposing one
%Use oppositecolour to generate threats and threat captures
function [potentialmoves,capt_index] = analyseboard(chessboard, piece_colour,num_moves,colour)

%Initialisation ----------------------------------------------------------
[p_x, p_y] = find(piece_colour == colour);
n_remaining = length(p_x);
potentialmoves = zeros(8,8);

%Loop to look at every piece's moves ------------------------------------
for i=1:n_remaining
    %Determines what piece is selected
    p_type = chessboard(p_x(i),p_y(i));

    %Based on the type of piece, its movement is calculated
    switch p_type
        case 1
            [move] = MovementPawn(chessboard,piece_colour,num_moves,p_x(i),p_y(i));
            %disp('Pawn');
        case 5
            [move] = MovementRook(chessboard,piece_colour,p_x(i),p_y(i));
            %disp('Rook');
        case 4
            [move] = MovementBishop(chessboard,piece_colour,p_x(i),p_y(i));
            %disp('Bishop');
        case 3
            [move] = MovementKnight(chessboard,piece_colour,p_x(i),p_y(i));
            %disp('Knight');
        case 9
            [move] = MovementQueen(chessboard,piece_colour,p_x(i),p_y(i));
            %disp('Queen');
        case 10
            [move] = MovementKing(chessboard,piece_colour,num_moves,potentialmoves,p_x(i),p_y(i));
            %disp('King');
    end

    %Sums up all possible moves of 1 colour.
    potentialmoves = potentialmoves+move;
end
%-------------------------------------------------------------------------
%                     Analysis of potentialmoves
%-------------------------------------------------------------------------
%-------------------Capture Analysis--------------------------------------
potentialcaptures = potentialmoves ~= 0 & chessboard~= 0;
capt_index = find(potentialcaptures==1);
%num_pot_capture = length(capt_index);
%capt_value_sum = sum(chessboard(capt_index));

%------------------ Moves Analysis ---------------------------------------
%nocapture = potentialmoves;
%nocapture(capt_index) = 0;
%num_moves_available = sum(sum(nocapture));

end
\end{lstlisting}

\subsubsection{KingCheck.m}
\begin{lstlisting}
%KingCheck Checks if the king is in check, checkmate or stalemate
%Colour in this case must be the current colour
%King Colour must be contrary to CAPT_INDEX & POTENTIAL MOVES
function [value]=KingCheck(chessboard,piece_colour,ownkingcolour, oppcolourcapt_index,oppcolourpotentialmoves)

%-------------------- King In Check --------------------------------------
king_index = find(chessboard == 10 & piece_colour == ownkingcolour);
kingincheck = ismember(king_index,oppcolourcapt_index);
if(kingincheck)
    value = 1;
%Otherwise not in check
else
    value = 0;
end
end

\end{lstlisting}

\subsubsection{checkmate.m}
\begin{lstlisting}
%Checkmate Determines if the currentboard is a checkmate state for
%specified colour
%Gives 1 for Checkmate, 0 for not checkmate
function [result]=checkmate(B,chessboard,piece_colour, num_moves)

if(mod(B.info.turn,2)==1)
    colour = 119;
    oppcolour = 98;
else
    colour = 98;
    oppcolour = 119;
end
result = 1;
%-------------------------------------------------------------------------
%        Loop that generates all possible moves
%-------------------------------------------------------------------------
[p_x,p_y] = find(piece_colour == colour);
n_remaining = length(p_x);
[potentialmoves] = analyseboard(chessboard, piece_colour,num_moves,oppcolour);

%In essence, we are going through each piece, looking at it's possible
%moves, make those possible moves, evaluate, save bestboard.
for i=1:n_remaining
    p_type = chessboard(p_x(i),p_y(i));
    switch p_type
        case 1
            [move] = MovementPawn(chessboard,piece_colour,num_moves,p_x(i),p_y(i));
        case 5
            [move] = MovementRook(chessboard,piece_colour,p_x(i),p_y(i)); 
        case 4
            [move] = MovementBishop(chessboard,piece_colour,p_x(i),p_y(i));
        case 3
            [move] = MovementKnight(chessboard,piece_colour,p_x(i),p_y(i));
        case 9
            [move] = MovementQueen(chessboard,piece_colour,p_x(i),p_y(i));
        case 10
            [move] = MovementKing(chessboard,piece_colour,num_moves,potentialmoves,p_x(i),p_y(i));
    end

%-------------------------------------------------------------------------
%            Individual Piece Moves That Generate New Game States
%                    Recursion is also added in each loop
%-------------------------------------------------------------------------
     [move_x,move_y] = find(move ~= 0);
     n_move = length(move_x);
%This loop generates all the game states from 1 piece
     for j = 1:n_move
         switch move(move_x(j),move_y(j))
             case 1
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickMovePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 2
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCapturePiece(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 3
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickEnpassant(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move, 0,1,move_x(j),move_y(j));
             case 4
                 [pchessboard, ppiece_colour, pnum_moves,kingincheck]=ClickCastling(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j));
             case 5
                 [pchessboard,ppiece_colour, pnum_moves,kingincheck]=ClickPawnPromo(0,0,p_x(i),p_y(i),B,piece_colour,chessboard,...
                     num_moves,0,move,0,1,move_x(j),move_y(j),'queen');
         end

         result = min(kingincheck, result);
         if result == 0
             break
         end

     end
     if result == 0
         break
     end
end
\end{lstlisting}

\subsubsection{readchessboard.m}
\begin{lstlisting}
%readchessboard takes in chessboard and creates B
function [B] = readchessboard(B,chessboard,piece_colour)

X = struct(NewPiece([]));
% build the initial board with everything non-playable at first
% add paddings to the non-playable areas of 4 squares and place pieces
for i=1:size(chessboard,1)+B.info.pad
    for j=1:size(chessboard,2)+B.info.pad
        X(i,j) = NewPiece([]);
    end
end

% now place pieces and playable areas
for i=1:size(chessboard,1)
    for j=1:size(chessboard,2)
        if chessboard(i,j) == 0
            pName = []; pColour = 0;
        else
            switch chessboard(i,j)
                case 1
                    pName = 'pawn';
                case 3
                    pName = 'knight';
                case 4
                    pName = 'bishop';
                case 5
                    pName = 'rook';
                case 9
                    pName = 'queen';
                case 10
                    pName = 'king';
            end

            switch piece_colour(i,j)
                case 119
                    pColour = 1;
                case 98
                    pColour = -1;
        end
        X(i+B.info.pad/2,j+B.info.pad/2) = NewPiece(pName,pColour);
    end
end
end

B.top = X;
end
\end{lstlisting}

\end{document}